---
title: "Homework 2"
output: github_document
---
```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

#### Reading and cleaning the data set

Read the Mr. Trash Wheel data set into R.

```{r read_trashwheel}
trashwheel_df =
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```


#### Reading and cleaning precipitation data for 2018

```{r read_precip_2018}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1,
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)
```


#### Reading and cleaning precipitation data for 2017
```{r read_precip_2017}
precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1,
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```


#### Combining annual precipitation
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

precip_month_name_df = 
left_join(precip_df, month_df, by = "month")
```


#### Describing the precipitation data



## Problem 2

#### Reading and cleaning transit data

```{r read_transit}
transit_df =
   read_csv(
     "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(entry = recode(entry, YES = TRUE, NO = FALSE))
 

  
sapply(transit_df, class)
```

#### Describing NYC Transit Data



#### How many distinct stations are there?
```{r distinct_stations}
distinct_stations = distinct(transit_df, station_name, line)
```
There are `r nrow(distinct_stations)` distinct stations in NYC.

#### How many stations are ADA compliant?
```{r ada_compliant}
ada_compliant = distinct(transit_df, station_name, line, ada) %>%
  filter(ada == TRUE)
```
There are `r nrow(ada_compliant)` ADA compliant stations in NYC.

#### What proportions of station entrances/exits without vending allow entrance?
```{r vending_entrance}
without_vending = filter(transit_df, vending == "NO")

allow_entrance = filter(without_vending, entry == TRUE)
```
The proportion of station entrances/exits without vending that allow entrance is `r round(nrow(allow_entrance)/nrow(without_vending), 3)`.
